%%%%%%%%%%%%%%%%
\begin{center}
\large \bf
ACKNOWLEDGMENTS
\end{center}
%%%%%%%%%%%%%%%%
\vspace{1.5cm}
%%%%%%%%%%%%%%%%
I feel truly grateful to have the opportunity to acknowledge the people who have been a part of my academic journey. First and foremost, I express my sincere respect and gratitude to the soldiers of India and to every taxpayer of the country, whose contributions have made it possible for me to pursue research in a safe and supportive environment.

I am deeply indebted to my Ph.D. supervisor, Prof. Debottam Das, and my co-supervisor, Retd. Prof. Suresh Kumar Patra, for their constant guidance, encouragement, and for giving me the opportunity to pursue research during my doctoral studies. Their mentorship has been invaluable in shaping both my scientific understanding and my approach to research.

I would also like to express my heartfelt gratitude to the members of my Doctoral Committee—Prof. Sanjib Kumar Agarwalla, Prof. Aruna Kumar Nayak, Prof. Manimala Mitra, and Prof. Prafulla Kumar Panda—for their insightful comments and constructive suggestions during my annual review seminars. Their valuable feedback helped me overcome several misconceptions and significantly broadened my understanding of physics.

I am fortunate to have worked with wonderful group members and collaborators. I sincerely thank Harishda for the many stimulating discussions during the early days of my Ph.D. A special word of thanks goes to Jeet, from whom I learned numerous practical research skills and with whom I found not only an excellent collaborator but also a true friend. Our countless conversations over coffee, ranging from physics to life and everything beyond, remain among my most cherished memories.

I would also like to thank the seniors at the Institute of Physics—Arpanda, Abhishekda, Manishda, Biswajitda, Kamleshda, and Kunduda—who never hesitated to extend their support, whether in academic matters or personal life. The late-night rooftop addas at the IOP New Hostel were filled with thoughtful conversations that taught me many lessons and, in many ways, shaped my perspective on life.

My heartfelt thanks also go to all my batchmates, each of whom brought a unique way of thinking about physics and enriched my learning experience. I am especially grateful to Rahul Puri for the many stimulating discussions on physics, to Subhalaxmi Rout for her friendship and support, and to Nevin Noble, whose enthusiasm and vibrant personality made every interaction memorable.

Finally, I owe my deepest gratitude to my family. My Maa and Bapi have been my greatest source of unconditional love, unwavering support, and endless encouragement throughout this uncertain and challenging journey. Their sacrifices and faith in me have been the foundation of everything I have achieved. I am equally grateful to my brother for always standing by my side and being a constant source of strength whenever I needed him.



\begin{comment}
cobut some members' contributions played exceptional roles in my Ph.D. journey. I want to offer my respect and gratitude to Dr. Masoom Singh. Without her help, I could not start my Ph.D. journey. At every step of my research work, the inspirations offered by Dr. Singh and Dr. Anil Kumar cannot be limited to words. I want to thank my collaborator, Dr. Suprabh Prakash, for giving remarkable inputs during our work. I want to thank Dr. Sudipta Das (Sudipta da) not only for helping me academically but also for showering selfless affection on me as an elder brother since the first day of my Ph.D. life and for supporting me at this difficult time of my life. I want to thank Mr. Pragyanprasu Swain for helping me both academically and non-academically. I want to thank Dr. Ashish Narang, Gopal, Sadashiv, Anuj, Krishnamoorthy, Sharmishtha, and Dr. Mehedi Masud for spending a joyful time with me. I am acknowledging Dr. Akhila Kumar Pradhan for giving suggestions on my thesis. I want to acknowledge all of my batchmates: Arpan, Chitrak, Ayesha, Mousam, Harish, Siddharth, Abhishek, Samir, Sandhya, and Pritam da. I am really grateful to the cordial, generous, and immediate help of Mr. Makrand Siddhabhatti (even beyond the office hours and at the weekends and holidays) and the entire team controlling the High-Performance Computing (HPC) facility of IOP, SAMKHYA, and the library facility. I am honored by getting the affiliation of the Department of Atomic Energy (DAE) and Homi Bhabha National Institute (HBNI), Mumbai. I am thankful to all the trees, mango garden, dogs, birds,and the serenity of the IOP campus.\\
    I want to thank my psychologist, Dr. Jayita Saha, for helping me handle the challenging situation during the last 6 years. I feel lucky to have some true friends like Abhishek Nandy, Soumyadipta Ray, Soumallya Banerjee, Sayan Jana, and Uma Das who always stood beside me. I want to thank Dr. Songshaptak De, Dr. Ramita Sarkar, Dr. Antara Dey, Dr. Sudipta Moshat, and Mr. Sunish Kumar Deb for supporting me during the hard time of my Ph.D. I extend my heartfelt love to all my seniors and juniors: Mr. Abhishek Bag, Dr. Ganesh Paul, Dr. Bibhabasu De, Dr. Saiyad Ashanujjaman, Dr. Debasish Saha, Dipak, Moonsun, Subhadip, and Alok Kumar. Mr. Debabrata Dey's academic suggestions helped me immensely, and I want to exceptionally thank him. I want to show my respect to Prof. Amitava Raychowdhury, Prof. Anindya Datta, Prof. Anirban Kundu, Prof. Indrajit Mitra, Prof. Jayashree Saha, and Prof. Sobhan Sounda. I will always be grateful to Prof. Poonam Mehta (JNU) for standing beside me at the dark time of my Ph.D. and helping me both academically and morally. Lastly, I bid my pranam to my physics teachers, Mr. Surojit Dhole and Mr. Vedakrishna Dey, and our Headmaster, Swami Kalyaneshananda, whose inspirations are still motivating me in carrying out the research on Physics and hopefully will illuminate the future path of my life.
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{1.0cm}
\begin{flushright}
	\textbf{(Debabrata Dey)}
\end{flushright}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
