%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\textbf{\large \underline{List of Publications for the thesis}}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{\large \underline{{Journal}}}\\
\textbf{\underline{Published}}
\begin{enumerate}
%%%%%%%%%%%%%%
\item \textit{Universality in spacetime $\omega$
	modes of quarkyonic stars} \\
\textbf{Debabrata Dey}, Jeet Amrit Pattanaik, R.N. Panda, S. K. Patra\\
\href{https://journals.aps.org/prd/abstract/10.1103/xs28-lt42}{\textit{Phys. Rev. D 113 (2026) 12, 123034,}} \href{https://arxiv.org/pdf/2111.11748}{arXiv:2602.22641v2 [astro-ph].}

%%%%%%%%%%%%%%%
\item  \textit{W-mode oscillation of neutron star in a new relativistic hybrid model}\\
\textbf{Debabrata Dey}, Jeet Amrit Pattanaik, R.N. Panda, S. K. Patra\\
\href{https://www.sciencedirect.com/science/article/pii/S0550321326002269}{\textit{Nucl. Phys. B 1029 (2026) 117519,}} \href{https://arxiv.org/abs/2509.01175}{arXiv: arXiv:2509.01175v1  [astro-ph].}
%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%
\item  \textit{f-mode oscillations of dark matter admixed quarkyonic neutron star}\\
\textbf{Debabrata Dey}, Jeet Amrit Pattanaik, M. Bhuyan, R.N. Panda, S. K. Patra\\
\href{https://iopscience.iop.org/article/10.1088/1475-7516/2025/08/003}{\textit{ JCAP 08 (2025) 003,}} \href{https://arxiv.org/abs/2412.06739}{ arXiv:2412.06739v1  [astro-ph].}
%%%%%%%%%%%%%%%
\item  \textit{Dark matter influence on quarkyonic stars: a relativistic mean field analysis}\\
\textbf{Debabrata Dey}, H. C. Das, Ankit Kumar, Jeet Amrit Pattanaik, R.N. Panda, S. K. Patra\\
\href{https://iopscience.iop.org/article/10.1088/1475-7516/2025/01/056}{\textit{JCAP 01 (2025) 056,}} \href{https://arxiv.org/abs/2401.021905}{ arXiv:2401.02190v2  [astro-ph].}
%%%%%%%%%%%%%%
%%%%%%%%%%%%%%
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{\underline{Work in progress}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{enumerate}
\item \textit{Core or Halo? Two-Fluid Analysis of Dark Matter-Admixed Quarkyonic Stars in the Multi-Messenger Era}\\ 
 Jeet Amrit Pattanaik, \textbf{Debabrata Dey}, M. Bhuyan, R.N. Panda, S. K. Patra \\
\textcolor{blue}{arXiv:2509.06684v1.}\\ 

\item \textit{Quarkyonic Model for Neutron Star Matter: A Relativistic Mean-Field Approach}\\ 
Ankit Kumar, \textbf{Debabrata Dey}, Shamim Haque, Ritam Mallick, S. K. Patra\\
\textcolor{blue}{ arXiv:2304.08223v2.}\\ 
\end{enumerate}  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{\large\underline{Conference Proceedings}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{enumerate}
%%%%%%%%%%%%%%%
	
\item \textit{Dark Matter Effects on f-mode Oscillations in Quarkyonic Stars}\\ 
\textbf{Debabrata Dey}, Jeet Amrit Pattanaik, M. Bhuyan, R.N. Panda, S. K. Patra
\\\href{ www.sympnp.org/proceedings}{DAE Symp.Nucl.Phys. 68 (2025) 765-766}
	
\item \textit{Influence of Dark Matter on Neutron Star Curvature within the Quarkyonic Model: A Relativistic Mean Field Approach}\\
\textbf{Debabrata Dey}, Jeet Amrit Pattanaik, M. Bhuyan, R.N. Panda, S. K. Patra\\\href{ www.sympnp.org/proceedings}{ DAE Symp.Nucl.Phys. 68 (2025) 763-764 }.
 
\item \textit\textit{Quarkyonic model for dark matter admixed neutron Star: A RMF perspective}\\
\textbf{Debabrata Dey}, H. C. Das, Ankit Kumar, Jeet Amrit Pattanaik, R.N. Panda, S. K. Patra\\
\href{www.sympnp.org/proceedings}{\textcolor{blue}{DAE Symp.Nucl.Phys. 67 (2024) 807-808.}}

\item \textit{Influence of Dark Matter on Neutron Star Curvature within the Quarkyonic Model: A Relativistic Mean Field Approach}\\
\textbf{Debabrata Dey}, Jeet Amrit Pattanaik, M. Bhuyan, R.N. Panda, S. K. Patra\\\href{ www.sympnp.org/proceedings}{ DAE Symp.Nucl.Phys. 68 (2025) 763-764 }.

\item \textit\textit{Tidal deformability of dark matter admixed neutron star within quarkyonic model in the conjunction of relativistic mean field formalism}\\
\textbf{Debabrata Dey}, H. C. Das, Ankit Kumar, Jeet Amrit Pattanaik, R.N. Panda, S. K. Patra\\
\href{www.sympnp.org/proceedings}{\textcolor{blue}{DAE Symp.Nucl.Phys. 67 (2024) 805-806.}}

\item \textit\textit{f-Mode Universal Relation of Dark Matter Admixed Quarkyonic Stars}\\
\textbf{Debabrata Dey}, Jeet Amrit Pattanaik, M. Bhuyan, R.N. Panda, S. K. Patra\\
\href{https://link.springer.com/chapter/10.1007/978-981-95-1513-4_51}{\textcolor{blue}{Springer Proc.Phys. 432 (2026) 231-234 .}}

\item \textit\textit{Quarkyonic model for dark matter admixed neutron Star: A RMF perspective}\\
\textbf{Debabrata Dey}, H. C. Das, Ankit Kumar, Jeet Amrit Pattanaik, R.N. Panda, S. K. Patra\\
\href{https://link.springer.com/chapter/10.1007/978-981-95-1513-4_51}{\textcolor{blue}{DAE Symp.Nucl.Phys. 67 (2024) 807-808 }}

\item \textit\textit{Tidal deformability of dark matter admixed neutron star within quarkyonic model in the conjunction of relativistic mean field formalism}\\
Jeet Amrit Pattnaik, R. N. Panda, \textbf{D. Dey}, Ankit Kumar, S. K. Patra, H. C. Das\\
\href{https://sympnp.org/proceedings/}{\textcolor{blue}{DAE Symp.Nucl.Phys. 67 (2024) 805-806.}}

%%%%%%%%%%%%%%%
\end{enumerate}
%\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{\large\underline{Talk/Poster}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{enumerate}

%%%%%%%%%%%%%%%
\item {\bf Talk:} \textcolor{blue}{\textit{f-Mode Universal Relation of Dark Matter Admixed Quarkyonic Stars}}\\
 26th DAE-BRNS High Energy Physics Symposium 2024,(In Person).
\vspace{0.5cm}	
\item {\bf Talk:} \textcolor{blue}{\it w-mode oscillation of quarkyonic star}\\
National Workshop on Recent Advances in Quantum Dynamics, Nuclear and Astrophysics, Institute Of Physics, Bhubaneswar; February 16-17, 2026 (In Person).
\vspace{0.5cm}
\item {\bf Poster:} \textcolor{blue}{\textit{Dark Matter Effects on f-mode Oscillations in Quarkyonic Stars}}\\
 68th DAE Symposium on Nuclear Physics, December 7-11, 2024 .
\vspace{0.5cm}	
\item {\bf Poster:} \textcolor{blue}{\textit{Influence of Dark Matter on Neutron Star Curvature within the Quarkyonic Model: A Relativistic Mean Field Approach}}\\
 68th DAE Symposium on Nuclear Physics, December 7-11, 2024.
\vspace{0.5cm}	
\item {\bf Poster:} \textcolor{blue}{\textit{Quarkyonic model for dark matter admixed neutron Star: A RMF perspective}}\\
67th DAE Symposium on Nuclear Physics, December 9-23, 2023.

\vspace{0.5cm}
\item {\bf Poster:} \textcolor{blue}{\textit{Tidal deformability of dark matter admixed neutron star within quarkyonic model in the conjunction of relativistic mean field formalism}}\\
: 67th DAE Symposium on Nuclear Physics, December 9-23, 2023.
\vspace{0.5cm}



%%%%%%%%%%%%%%%
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{0.5cm}
\begin{flushright}
	\textbf{(Debabrata Dey)}
\end{flushright}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%